---
title: "Practical Machine Learning - Prediction Assignment Writeup"
---

The goal of your project is to predict the manner in which users did the exercise.This is the "classe" variable in the training set. 

Based on the data recorded from several people using Activity Monitors, we will try to construct a macine learning algorithm to predict future outcomes on the test data. I am using Random Forest Modeling, and got over 99% accuracy on my training set of 70% of the total data. All my submitted test cases were succesful.

### Data Availability
The training data for this project are available here: 
  https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data are available here: 
  https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

### Loading and cleaning Data

```{r}
library(caret)
testing <- read.csv("pml-testing.csv",na.strings=c("NA",""))
training <- read.csv("pml-training.csv",na.strings=c("NA",""))
```

Discarding columns with not enough data

```{r fig.width=7, fig.height=6}
var <- names(training)[apply(training,2,function(x) table(is.na(x))[1]==19622)]   
training02<- training[,var]
testing02 <- testing[,c(var[-length(var)],names(testing)[length(testing)])] 

unneedolIndex <- grep("new_window|num_window|X|timestamp|user_name",colnames(training02))
training03 <- training02[,-c(unneedolIndex)]
testing03  <- testing02[,-c(unneedolIndex)]
```


### Building data sets for training and cross validation.

##### Using 70% for training and 30% for Cross Validation. None generated for testing since that set is already provided.

```{r}
yy = createDataPartition(training03$classe, p = 0.7, list = FALSE)
trainPart = training03[ yy,]
testPart = training03[-yy,]
```

#### Training...
randomForest is used as it is one of the best classification 

```{r}
library(randomForest)
resultModel <- randomForest(classe ~ .,data = trainPart)
resultModel
```

#### Calculation the errors using the Cross Validation Set.

###### cross validation

```{r}
aTestingLabelsPred <- predict(resultModel, newdata = testPart)
aConfusionMatrix <- confusionMatrix(aTestingLabelsPred, testPart$classe)
aConfusionMatrix
```

###### Out-of-sample error

```{r}
(1 - aConfusionMatrix$overall["Accuracy"])[[1]]
```

#### Generating data for the prediction vector for the Assigment Submission

###### With the prediction model, we use it to predict 20 different test cases : 
  
```{r}
answers <- predict(resultModel,testing03)

pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}
pml_write_files(answers)
```


#### Result 

```{r}
answers
```